Computationally Efficient High-Dimensional Regularized Discriminant Analysis

### 2017-01-22 reports/paper-hdrda.Rnw
Revamped the abstract with Phil. Much better.

We develop a new and computationally efficient regularized discriminant analysis (RDA)
classifier for the small-sample, high-dimensional setting.  We are able
to interpret the biased covariance matrix estimator employed in the HDRDA
classifier as one that is estimated from all training observations regardless of
class. Moreover, our classifier also differs from
\citeauthor{Friedman:1989tm}'s in that we show our classifier is equivalent to
a classifier in a reduced feature space with dimension approximately equal to the
training sample size. As a result, the matrix operations used in the HDRDA
classifier are computationally linear in the number of features, making HDRDA
well-suited for high-dimensional classification in practice. We demonstrate that HDRDA is often superior to several
sparse and regularized classifiers in terms of classification accuracy with
three artificial and six real high-dimensional data sets.  Also, timing comparisons
between our HDRDA implementation in the {\tt sparsediscrim} R package and the
standard RDA formulation in the {\tt klaR} R package demonstrate that as the
number of features increases, the computational runtime of HDRDA is drastically
smaller than RDA.

@@

The choice of $\alpha_k$ in \eqref{eq:hdrda-cov} is one of convenience and
allows the flexibility of various covariance-matrix estimators proposed in the
literature. In practice, we generally are not interested in estimating
$\alpha_k$ because the estimation of $K$ additional tuning parameters via
cross-validation is counterproductive to our goal of improving the \emph{RDA}
classifier's computational efficiency. By setting $\alpha_k$ equal to an
appropriate constant, the \emph{HDRDA} covariance-matrix estimator includes or
resembles a large family of estimators. Notice that if $\alpha_k = 1$ and
$\lambda = 1$, \eqref{eq:hdrda-cov} is equivalent to the standard ridge-like
covariance-matrix estimator in \eqref{eq:ridge-estimator}, and various
estimators proposed in the literature can be obtained when one selects $\gamma$
accordingly. For instance, with $\gamma = \tr\{ \widehat{\bm \Sigma} \} /
\min(N, p)$, we obtain the estimator from \cite{Srivastava:2007ww}.

>>>>

Regularized discriminant analysis (\emph{RDA}) proposed by
\cite{Friedman:1989tm} is a widely popular classifier that lacks
interpretability and is impractical for high-dimensional data sets.  Here, we
present an interpretable and computationally efficient classifier called
high-dimensional \emph{RDA} (\emph{HDRDA}), designed for the small-sample,
high-dimensional setting. For \emph{HDRDA}, we show that each training
observation regardless of class contributes to the class covariance matrix,
resulting in interpretable estimator that \emph{borrows} from the pooled sample
covariance matrix. Moreover, we show that \emph{HDRDA} is equivalent to a
classifier in a reduced feature space with dimension approximately equal to the
training sample size. As a result, the matrix operations employed by
\emph{HDRDA} are computationally linear in the number of features, making the
classifier well-suited for high-dimensional classification in practice. We
demonstrate that \emph{HDRDA} is often superior to several sparse and
regularized classifiers in terms of classification accuracy with three
artificial and six real high-dimensional data sets. Also, timing comparisons
between our \emph{HDRDA} implementation in the {\tt sparsediscrim} R package
and the standard \emph{RDA} formulation in the {\tt klaR} R package demonstrate
that as the number of features increases, the computational runtime of
\emph{HDRDA} is drastically smaller than \emph{RDA}.

@@

The choice of $\alpha_k$ in \eqref{eq:hdrda-cov} is one of convenience and
allows the flexibility of various covariance-matrix estimators proposed in the
literature. In practice, we generally are not interested in estimating
$\alpha_k$ because the estimation of $K$ additional tuning parameters via
cross-validation is counterproductive to our goal of computational efficiency.
By setting $\alpha_k$ equal to an appropriate constant, the \emph{HDRDA} covariance-matrix estimator includes or
resembles a large family of estimators. Notice that if $\alpha_k = 1$ and
$\lambda = 1$, \eqref{eq:hdrda-cov} is equivalent to the standard ridge-like
covariance-matrix estimator in \eqref{eq:ridge-estimator}. Other
estimators proposed in the literature can be obtained when one selects $\gamma$
accordingly. For instance, with $\gamma = \tr\{ \widehat{\bm \Sigma} \} /
\min(N, p)$, we obtain the estimator from \cite{Srivastava:2007ww}.
